### 2014-06-01 patchy-selection-paper.tex
note
edited applications and altered scaled rec rate to reflect ratio of areas

As our goal here to describe the rough genomic length scale, 
based on the above, we view 
the lineage is simply bouncing around in the patch backward in time and we propose a
simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (eqn. \ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e. were both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
is roughly $2(\sigma/\sqrt{s})A^{-1/d}$, and in these regions $p(x)
\approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is, taking in this region, roughly 
\begin{equation}
r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
\end{equation}
We can use this to first of all think about the length of haplotype
associated with the $B$ allele after a given length of time $\tau$,
this could the the time since the sweep within a patch due to a
mutation or migration of a beneficial allele. If the recombination
rate per base is $r_{BP}$ then the length of haplotype associated with
the $B$ allele is
\begin{equation}
\frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
\end{equation} 
note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$
background.

@@

The mutational target size $\mu$ for the trait is unclear. 
While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
it is likely that not all of these changes are needed for a population to begin to  adapt. 
Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
have been identified in closely related species and more broadly across vertebrates \citep{REF}.
To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
where the adaptive mtation has to occur at a single base pair; 
and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs,
e.g.\ any of the coding bases of a typical gene.  
Finally, we set $A=100\text{km}^2$.
Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
of parallel mutation on the distance between lava flow patches using these parameters. 
As can be seen parallel mutation should become
likely over a scale of tens to a few hundred kilometers between patches.

@@

@@

small-scale fluctuations

when landscape is not made of discrete (circular) patches

>>>>

As our goal here to describe the rough genomic length scale, 
based on the above, we view 
the lineage is simply bouncing around in the patch backward in time and we propose a
simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (eqn. \ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e. were both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
multiplied by $p(x)$ in these intermediate regions, where we assume
$p(x) \approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is, taking in this region, roughly 
\begin{align}
&=
    \begin{cases}
    r \frac{\frac{\sigma}{\sqrt{s}} }{\frac{\sigma}{\sqrt{s}} + A }    \qquad & d=1 \\
   r \frac{\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
  \sqrt{\frac{A}{\pi}} \right)}{A+\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
  \sqrt{\frac{A}{\pi}} \right) }    \qquad & d=2 \\  
\end{cases}
\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
\end{align}
\gc{Altered the above.}
We can use this to first of all think about the length of haplotype
associated with the $B$ allele after a given length of time $\tau$,
this could the the time since the sweep within a patch due to a
mutation or migration of a beneficial allele. If the recombination
rate per base is $r_{BP}$ then the length of haplotype associated with
the $B$ allele is
\begin{equation}
\frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
\end{equation} 
note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$
background.

@@

The mutational target size $\mu$ for the trait is unclear. 
While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
it is likely that not all of these changes are needed for a population to begin to  adapt. 
Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
have been identified in closely related species and more broadly across vertebrates \citep{REF}.
To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
where the adaptive mutation has to occur at a single base pair; 
and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
being able to arise at any of a thousand base pairs.  
Finally, we set $A=100\text{km}^2$.
Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
of parallel mutation on the distance between lava flow patches using these parameters. 
As can be seen parallel mutation should become
likely over a scale of tens to a few hundred kilometers between patches.

Givn the large selection coefficient associated with the dark allele
on the dark substrate we expect the initial sweep associated with
either a new mutation or migrant allele to be large ($s/\log(\rho A
s)$). We can use eqn \eqref{eqn:haplotype_length} to calculate how long a
founding haplotype we initially expect to find shared between populations
this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
quite long between geographically close patches (tens of km). However,
for our wider cline width ($\sigma/\sqrt{s_m} =30km$) adaptation by
migration can still be likely for patches 100km apart but its shared
basis may be hard to detect as the length of shared haplotype can be
quite short.

Given the relatively small size of the dark substrate patches where
$A$ is often $<10km$ we should expect the haplotype to be rapidly
whittled down within patches (eqn. \ref{})

@@

recessive alleles will have different dynamics. Actually not really
obvious what will happen here, cline width will be much wider but
prob. of taking off will go as the square of the occupancy density.

@@

small-scale fluctuations
when landscape is not made of discrete (circular) patches
